---
title: "CDA Homework 7"
author: "Ryan Tracy"
date: "5/13/2019"
output: pdf_document
---

```{r, include = FALSE}
library(lme4)
library(lmerTest)
library(car)
```


#8.2

##a.
```{r}
(zTest <- (125 - 2) / (sqrt(125 + 2)))
chSq <- zTest^2
1-pchisq(chSq, 1)
```

##b.
```{r}
#sample proportion for heaven
(heaven <- (833+125) / (833+2+160+125))
#sample proportion for hell
(hell <- (833 + 2) / (833+2+125+160))
#difference
(diff <- heaven - hell)
#std error
(stErr <- (1/1120)*sqrt((125+2) - ((125-2)^2/1120)))
#90% CIs
diff + (1.645*stErr)
diff - (1.645*stErr)
```
The sample proportions are .855 for heaven and .746 for hell, $90\% \ CI [.094, .125]$.


#4.20
```{r}
clinical <- as.data.frame(rep(c(1:8), each = 2)); names(clinical)[1] <- "center"
clinical$treatment <- rep(c("drug", "control"), each = 1)
clinical$success <- c(11,10,16,22,14,7,2,1,6,0,1,0,1,1,4,6)
clinical$failure <- c(25,27,4,10,5,12,14,16,11,12,10,10,4,8,2,1)
```

##a.
```{r}
mod <- glm(success / (success + failure) ~ treatment, weights = success + failure, family = binomial, data = clinical)
summary(mod)
exp(mod$coefficients)
exp(confint(mod))
```
The logistic regression model revealed no significant effect for treatment, $b = .40,\ SE = .25,\ z = 1.61,\ p = .11$. While the drug may be associated with a .40 increase in the probability of sucesss, the fact that the effect is nonsignificant makes this a useless conclusion.

##b.
```{r}
clinicTable <- matrix(c(55, 47, 75, 96), nrow = 2, byrow = T)
rownames(clinicTable) <- c("drug", "control")
colnames(clinicTable) <- c("success", "failure")
clinicDat <- array(c(11,25,
                     10,27,
                     16,4,
                     22,10,
                     14,5,
                     7,12,
                     2,14,
                     1,16,
                     6,11,
                     0,12,
                     1,10,
                     0,10,
                     1,4,
                     1,8,
                     4,2,
                     6,1),
                   dim = c(2,2,8),
                   dimnames = list(
                     response = c("success", "failure"),
                     treatment = c("drug", "control"),
                     clinic = c("1", "2", "3", "4", "5", "6", "7", "8")))
clinicDat
mantelhaen.test(clinicDat)
```
The CMH test revealed that there was conditional dependence between the nominal variables in the clinical dataset, $\chi^2(1) = 5.67,\ p = .02$.

##"c."
```{r}
mod1 <- glmer(success / (success + failure) ~ treatment + (1|center), family = binomial, weights = success + failure, data = clinical)
Anova(mod1)
summary(mod1)
confint(mod1)
exp(.7379); exp(c(.1580966, 1.33952684))
```
This data is actually hierarchical in nature, so a random intercept of center may clarify whether or not the treatment does have an effect. Significance tests revealed that the effect of clinic did vary at random, $SD = 1.39,\ 95\%\ CI[.80, 2.69]$. This model also revealed that the effect of treatment was significant, $b = .74,\ SE = .30,\ z = 2.48,\ p = .01$. Patients who received the drug were 2.09 times as likely to be cured than those who did not, $OR = 2.09,\ 95\%\ CI[1.17, 3.82]$. This is a higher than the OR for the model that did not account for the random intercept of clinic, which was also not significant ($OR = 1.50,\ 95\%\ CI[.92, 2.46]$).


