---
title: "CatData HW3"
author: "Matthew Vanaman"
date: "03/11/19"
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
```

A2.2, A2.5, A2.13, A2.21, A2.36. (3 points each)

A2.33 (5 points)

*All code and work are shown in the appendix.*

# 2.2

## (a)
**Answer:**\newline
Sensitivity denotes the conditional probability of not having the disease given a negative diagnosis. In a sense, this captures the test's ability to accurately diagnose a patient that does not have the disease. We therefore know that the conditional probability of not having the disease given a patient were given a positive diagnosis is 1 - specificity, as this captures the noise of the test. A good test would want to show that 1 - specificity yields a large probability, as an effective test will have a large probability of correctly diagnosing a patient that does not have the disease. 

**Work:**\newline
