---
title: "HW 4"
author: "Matthew Vanaman"
date: "03/11/19"
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
```

4.1, 4.2, 4.3, 4.4, 4.5, 4.6

# 4.1

**Answer:**\newline
As n approaches infinity, the shape would approach a circle. 

# 4.2

## (a)
**Answer:**\newline
32

**Work:**\newline

$$\lim_{x \to 5} 2x^2 - 5x + 7,$$\newline
$$2(5^2) - 5(5) + 7,$$\newline
$$32.$$

```{r}
# chck work
2*(5)^2 - (5*5) + 7
```

## (b)
**Answer:**\newline
Limit is 0. Approaches but never quite gets there from the left nor right.

**Work:**\newline
$$\lim_{y \to \infty} \frac {1} {y^6}$$
```{r, results='hold'}
# check work
1 / 2^6
1 / 4^6
1 / 20^6
1 / -2^6
1 / -4^6
1 / -20^6
```

## (c)
**Answer:**\newline
Positive infinity. The larger the absolute value of z, the larger the quotient. 

**Work:**\newline
$$\lim_{z \to 0} \frac {1} {z^6}$$
```{r, results='hold'}
# check work
1 / 1^6
1 / 0.5^6
1 / 0.01^6
1 / (-1)^6
1 / (-0.5)^6
1 / (-0.01)^6
```


## (d)
**Answer:**\newline
The limit is 0.

**Work:**\newline
```{r, results='hold'}
# check work
2*2 / 6 * (2^2)
2*5 / 6 * (5^2)
2*10 / 6 * (10^2)
2*1 / 6 * (1^2)
2*0.1 / 6 * (0.1^2)
2*0.0001 / 6 * (0.001^2)
```
Could also solve this way:\newline
$$\lim_{x \to 0} \frac {2x + 3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2x} {5x^2}  +\frac {3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2} {5x}  +\frac {3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2} {5x}  + \lim_{x \to 0} \frac {3} {5x^2},$$\newline
$$\frac {2} {5} \lim_{x \to 0} \frac {1} {x}  + \frac {3} {5} \lim_{x \to 0} \frac {1} {x^2},$$\newline
$$\frac {2} {5} \lim_{x \to 0} \frac {1} {x}  + \frac {3} {5} \lim_{x \to 0} \frac {1} {x^2},$$\newline
The limits to $\frac {1} {x}$ and $\frac {1} {x^2}$ are 0.


## (e)
**Answer:**\newline


**Work:**\newline


## (f)
**Answer:**\newline


**Work:**\newline

## (g)
**Answer:**\newline


**Work:**\newline

## (h)
**Answer:**\newline


**Work:**\newline

## (i)
**Answer:**\newline


**Work:**\newline
