---
title: "HW 4"
author: "Matthew Vanaman"
date: "03/11/19"
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

4.5, 4.6

# 4.1

**Answer:**\newline
As n approaches infinity, the shape would approach a circle. 

# 4.2

## (a)
**Answer:**\newline
32

**Work:**\newline

$$\lim_{x \to 5} 2x^2 - 5x + 7,$$\newline
$$2(5^2) - 5(5) + 7,$$\newline
$$32.$$

```{r}
# chck work
2*(5)^2 - (5*5) + 7
```

## (b)
**Answer:**\newline
Limit is 0. Approaches but never quite gets there from the left nor right.

**Work:**\newline
$$\lim_{y \to \infty} \frac {1} {y^6}$$
```{r, results='hold'}
# check work
1 / 2^6
1 / 4^6
1 / 20^6
1 / -2^6
1 / -4^6
1 / -20^6
```

## (c)
**Answer:**\newline
Positive infinity. The larger the absolute value of z, the larger the quotient. 

**Work:**\newline
$$\lim_{z \to 0} \frac {1} {z^6}$$
```{r, results='hold'}
# check work
1 / 1^6
1 / 0.5^6
1 / 0.01^6
1 / (-1)^6
1 / (-0.5)^6
1 / (-0.01)^6
```


## (d)
**Answer:**\newline
The limit is 0.

**Work:**\newline
```{r, results='hold'}
# check work
2*2 / 6 * (2^2)
2*5 / 6 * (5^2)
2*10 / 6 * (10^2)
2*1 / 6 * (1^2)
2*0.1 / 6 * (0.1^2)
2*0.0001 / 6 * (0.001^2)
```
Could also solve this way:\newline
$$\lim_{x \to 0} \frac {2x + 3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2x} {5x^2}  +\frac {3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2} {5x}  +\frac {3} {5x^2},$$\newline
$$\lim_{x \to 0} \frac {2} {5x}  + \lim_{x \to 0} \frac {3} {5x^2},$$\newline
$$\frac {2} {5} \lim_{x \to 0} \frac {1} {x}  + \frac {3} {5} \lim_{x \to 0} \frac {1} {x^2},$$\newline
$$\frac {2} {5} \lim_{x \to 0} \frac {1} {x}  + \frac {3} {5} \lim_{x \to 0} \frac {1} {x^2},$$\newline
The limits to $\frac {1} {x}$ and $\frac {1} {x^2}$ are 0.


## (e)
**Answer:**\newline
$\frac {3} {5}$


**Work:**\newline
We know that the limit of $\frac {1} {x^y} = 0$, so we can reduce many of the terms of the expression to 0s by making them into the form of $\frac {1} {x^y} = 0$.

$$ \lim_{y \to \infty} \frac {3y^7 + 4y^6 - 2y^5 - 8y^3 - 7y + 1} {2y^7 + y^3 - 8} \times \frac { \frac {1} {y^7} } {\frac {1} {y^7} },$$\newline
$$ = \lim_{y \to \infty} \frac {3y^7 + \frac {4} {y} - \frac {2} {y^2} - \frac {8} {y^4} - \frac {7} {y^6} + \frac {1} {y^7}} {2 + \frac {1} {4y^4} - \frac {8} {y^7}},$$\newline
$$ = \frac {3} {5}.$$\newline


## (f)
**Answer:**\newline
The limit is 1.

**Work:**\newline

Do some factoring
$$\lim_{z \to 3} \frac {z^2 - 5z + 6} {z - 3},$$\newline


$$=\lim_{z \to 3} \frac {(z-2)(z-3)} {z - 3},$$\newline
$$ = \lim_{z \to 3} z-2 = 1$$\newline

```{python}
# check work
import sympy
from sympy import *
x, y, z = symbols('x y z')
init_printing(use_unicode=True)
expr = Limit((z - 2)*(z - 3) / (z - 3), z, 3)
# find the limit
print(expr.doit())
```

## (g)
**Answer:**\newline
$\infty$

**Work:**\newline
Need limit from the right:
```{r, results = 'hold'}
1 / (5.1 - 5)
1 / (5.01 - 5)
1 / (5.00001 - 5)
```

## (h)
**Answer:**\newline
The limit from the right is $\infty$ while the limit from the left is $-\infty$, therefore there is no limit. 

**Work:**\newline
Need limit from the right:
```{r, results = 'hold'}
1 / (7.1 - 7)
1 / (7.01 - 7)
1 / (7.00001 - 7)
```
and left:
```{r, results = 'hold'}
1 / (6.9 - 7)
1 / (6.99 - 7)
1 / (6.99999- 7)
```


## (i)
**Answer:**\newline
$\approx 7.388611$


**Work:**\newline
Aha! This is just Euler's number, which is equal to 2.7182... Therefore:

$$\lim_{z \to \infty} \bigg( 1 + \dfrac {1} {z} \bigg)^{2z},$$
Becaause the exponenents are mutliplicative, put the z on the inside to get euler's number:
$$= \lim_{z \to \infty} \Bigg( \bigg( 1 + \dfrac {1} {z} \bigg)^z \Bigg)^2,$$
$$= \lim_{z \to \infty} (e)^2$$
$$= \lim_{z \to \infty} (2.7182)^2$$
$$\approx 7.388611$$

```{r}
# calculations
2.7182^2
```


# 4.3

## (a)

**Answer:**\newline
When Barack Obama says that the deficit is changing at a certain pace, we can think of that pace as being the slope of the tangent line at a given point along the x- and y-axes, with the x-axis representing time and the y-axis representing debt. We can really only approximate the rate at which debt is falling with the tangent line (i.e. instantaneous rate of change). The tangent line is just the secant line between two points very very close together along the line (line being the deficit throughout the years), and we could achieve greater and greater resolution by taking the derivative as many times as needed. Given the function $f(t)$ represents the national debt for a given year ($t$), then in our case we could get the average rate of change for that year by taking the first derivative: $f'(t)$. However, that's not enough resolution: to make a claim about the *current rate* at which the deficit is falling, you'd need the instantaneous rate of change. To get that, you'd take the second derivative, $f''(t)$. This gives you the value of the tangent line for a given t. We also know that this has to be a negative number because the deficit is going down (i.e. tangent line is a negative slope, indicating that the slopoe of the overall curve is decreasing).

## (b)

**Answer:**\newline
This trend isn't accurate, but is true in the sense that 2013 would indeed be the year with the fastest falling deficit:
```{r}
sigmoid = function(x) {
   1 / (1 + exp(-x))
}
x <- seq(-5, 5, 0.01)
p1 <- ggplot(data_frame())

p1 + stat_function(aes(-4:3), fun = sigmoid) + ylab("Debt") + 
  theme(
  axis.text.y = element_blank(),
  axis.ticks = element_blank())+
  scale_x_continuous(name = "Year", breaks = c(-4, -2, 0, 2), 
                     labels = c("1953", '1973', '1993', '2013'), 
                     limits = c(-4, 2))
```

# 4.4
A rough approximation:
```{r, warning = FALSE}
x <- seq(-5, 5, 0.001)
dat <- data.frame(x,y=(2*x)^2)
f <- function(x) (x*2)^2
ggplot(dat, aes(x,y)) + 
    geom_line()  + 
  theme(
  axis.ticks = element_blank())+
  scale_x_continuous(name = "x", breaks = c(
    -1, 0, 1), 
    labels = c(-4, 0, 4),
    limits = c(-2, 2)) +
    scale_y_continuous(name = "y", breaks = c(
    -1, 0, 1, 2, 3, 4, 5), 
    labels = c(-2, -1, 0,1, 2, 3, 4),
    limits = c(-1, 5))
```




# 4.5

## 4.5(a)

**Answer:**\newline
$= 6x^{5} + 25x^{4} - 4x.$


**Work:**\newline
$$\textrm {if} \ f(x) = x^6 + 5x^5 - 2x^2 + 8, \textrm {then}$$
$$f'(x) = \dfrac {d} {dx} \Big( x^6 + 5x^5 - 2x^2 + 8 \Big),$$
Use the addition/subtraction shortcut:
$$= \dfrac {d} {dx}(x^6) + \dfrac {d}{dx} (5x^5) - \dfrac {d} {dx} (2x^2) + \dfrac {d} {dx} (8),$$
$$= 6x^{6-1} + 5 (5x^{5-1}) - 2 (2x^{2-1}) + 0,$$
$$= 6x^{5} + 25x^{4} - 4x.$$

## 4.5(b)

**Answer:**\newline
$ = 3e^{y}  - \dfrac {1} {2\sqrt{y}}.$

**Work:**\newline

$$\textrm{if} \ g(y) = 3e^y - \sqrt{y}, \ \textrm{then}$$
$$g'(y) = \dfrac {d} {dy} \Big( 3e^y - \sqrt{y} \Big),$$
$$ = 3 \dfrac {d} {dy} (e^y)  - \dfrac {d} {dy} (\sqrt{y}),$$
$$ = 3\dfrac {d} {dy} (e^y)  - \dfrac {1} {2\sqrt{y}},$$
Assuming the $e$ here is Euler's number and not a variable like $x$ or $y$ or another constant like $a$, then $\dfrac {d} {dy} (e^y)$ is just $e^y$:
$$ = 3e^{y}  - \dfrac {1} {2\sqrt{y}}.$$

## 4.5(c)

**Answer:**\newline
$\dfrac {1} {z} - \dfrac {1} {z^{2}}  + \ln{(3)}3^z.$

**Work:**\newline

$$\textrm{if} \ h(z) = ln(z) + \dfrac {1} {z} + 3^z, \ \textrm{then}$$
$$h'(z) = \dfrac {d} {dz} \big( \textrm{ln}(z) \big) + \dfrac {d} {dz} \big( \dfrac {1} {z} \big) + \dfrac {d} {dz}(3^z),$$
$$\dfrac {1} {z} + \dfrac {d} {dz} \big( \dfrac {1} {z} \big) + \dfrac {d} {dz}(3^z),$$
$$\dfrac {1} {z} + \dfrac {-1} {z^{1+1}}  + \dfrac {d} {dz}(3^z),$$
$$\dfrac {1} {z} - \dfrac {1} {z^{2}}  + \dfrac {d} {dz}(3^z),$$
$$\dfrac {1} {z} - \dfrac {1} {z^{2}}  + \ln{(3)}3^z.$$


## 4.5(d)


**Answer:**\newline
$\Big( (x + 3)^7 \Big) \Big(12x^3 - 4x \Big) + \Big(7(x+3)^6 \Big) \Big(3x^4 - 2x^2 - 8\Big).$

**Work:**\newline
$$j(x) = (x + 3)^7(3x^4 - 2x^2 - 8).$$
$$j'(x) = \dfrac {d} {dx}(x + 3)^7 \dfrac {d} {dx}(3x^4 - 2x^2 - 8).$$
Would take too long to FOIL. To get derivative of first half, break it into layers:
$$A^7 = (x + 3)^7, \ A = x + 3,$$
Take derivative of $A^7$ with power rule:
$$h'(x) = \dfrac {d} {dx}(A^7) = 7(A^{7-1}) = 7A^6.$$
Lastly, substitute for A:
$$7A^6 = 7(x+3)^6.$$
For second half:
$$\dfrac {d} {dx}(3x^4 - 2x^2 - 8),$$
$$= \dfrac {d} {dx}(3x^4) - \dfrac {d} {dx}(2x^2) - \dfrac {d} {dx}(8),$$
$$= 3 \dfrac {d} {dx}(x^4) - 2 \dfrac {d} {dx}(x^2) - 0,$$
$$= 3 (4x^{4-1}) - 2 (2x^{2-1}),$$
 $$= 12x^{4-1} - 4x^{2-1},$$
 $$= 12x^3 - 4x.$$
Plugging the derivatives back into the function:
$$j'(x) = \dfrac {d} {dx}(x + 3)^7 \dfrac {d} {dx}(3x^4 - 2x^2 - 8),$$
$$= \Big( (x + 3)^7 \Big) \Big(12x^3 - 4x \Big) + \Big(7(x+3)^6 \Big) \Big(3x^4 - 2x^2 - 8\Big).$$
