---
title: "HW 4"
author: "Matthew Vanaman"
date: "03/11/19"
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
```

4.1, 4.2, 4.3, 4.4, 4.5, 4.6

# 4.1

**Answer:**\newline
As n approaches infinity, the shape would approach a circle. 

# 4.2

## (a)
**Answer:**\newline


**Work:**\newline

$$\lim_{x \to 5} 2x^2 - 5x + 7,$$\newline
$$2(5^2) - 5(5) + 7,$$\newline
$$32.$$

```{r}
# chck work
2*(5)^2 - (5*5) + 7
```

## (b)
**Answer:**\newline


**Work:**\newline

## (c)
**Answer:**\newline


**Work:**\newline

## (d)
**Answer:**\newline


**Work:**\newline

## (e)
**Answer:**\newline


**Work:**\newline

## (f)
**Answer:**\newline


**Work:**\newline

## (g)
**Answer:**\newline


**Work:**\newline

## (h)
**Answer:**\newline


**Work:**\newline

## (i)
**Answer:**\newline


**Work:**\newline
