---
title: "CatData HW 4"
author: "Matthew Vanaman"
date: "03/18/19"
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```


3.9, 3.15 (do not verify for 3.15d)

# 3.9
## (a)
**Answer:** $\text{logit} [\hat P(Y = 1)]= 0.0532x -3.5561$. In this example, x is income, Y is possession of travel card  (0 = does not possess, 1 = does possess).

## (b)
**Answer:** $\hat \beta$ means, in this case, that as income increases, the probability of possessing a travel card increases (because $\hat \beta$ is positive).

## (c)
**Answer:** First, we know that $\hat \pi$ denotes $P(Y=1)$, and that the logit function is $\text {log} \Bigg[ \dfrac {P(Y=1)} {1-P(Y=1)} \Bigg]$. When you substitute 0.50 in for $P(Y=1)$, you get 0:
```{r}
log(0.50 / (1 - 0.50))
```

To show when the etsimated $\pi$ is 0.5, we just need to solve for x:
```{python, echo=FALSE}
# Check Work
import sympy
from sympy import *
from sympy import Symbol
x = Symbol('x')
# Solve
print("solveset(Eq(-3.556 + 0.0532*x)), 0.50")
print(solveset(Eq(-3.556 + 0.0532*x)), 0.50)
```
Therefore, when x = 66.84, $\hat \pi = 0.50$. According to the model, there is a 50/50 shot of owning a travel card for those who make around 66.86 million lira.





# Question 3, AD Data

```{r}
q3Data <- as.data.frame(matrix(c(730,130,860,
                    100, 40, 140,
                    830,170,1000), ncol = 3))
colnames(q3Data) <- c("NLD", 'LD', 'Margin')
rownames(q3Data) <- c('NAD', 'AD','Margin')
knitr::kable(q3Data)
```

